Ludwig User Tutorial

### 2018-10-05 docs/tutorial/tutorial.tex

The system requirements are an up-to-date version of a C compiler as well as working installations 
of a text editor like \texttt{vi} or \texttt{emacs}, the visualisation tool
ParaView (\hyperref[ParaView]{https://www.paraview.org}), the MPI-library 
for compiling the parallel version of the code and \texttt{gnuplot} for simple visualisations.

@@

The \texttt{Ludwig} project is currently hosted at our central repository at GitHub:\\
(\hyperref[GitHubLudwig]{https://github.com/ludwig-cf/ludwig}).\\

\smallskip
You can either download or clone the repository. 
The following command clones the repository into \texttt{ludwig} and requires a working 
git installation on your local system:

@@

\begin{lstlisting}
$ ls ludwig
branches  tags  trunk
\end{lstlisting}
which is the usual svn layout for trunk, branches and tagged versions.

A full documentation of \texttt{Ludwig} can be obtained by running 
the command \texttt{make pdf} twice in the \texttt{/trunk/docs} directory:

\begin{lstlisting}[style=terminalverbatim]
$ cd /ludwig/trunk/docs/
$ make pdf
\end{lstlisting}

@@

\begin{enumerate}
\item Go to the directory \texttt{/ludwig/trunk/}: \\
\begin{lstlisting}
$ cd ludwig/trunk
\end{lstlisting}
\item Copy the configuration file \texttt{lunix-gcc-default.mk} to the \texttt{/trunk} directory 
and rename to \texttt{config.mk} (both steps exectued below at once): \\
\begin{lstlisting}
$ cp config/lunix-gcc-default.mk ./config.mk
\end{lstlisting}
Note that you may need to specify the path to the MPI wrapper compiler in this file.
\item Issue \texttt{make} in \texttt{/targetDP}: \\
\begin{lstlisting}
$ cd targetDP/
$ make 
\end{lstlisting}
This creates dummy functions for calls to the targetDP library.
\item Issue \texttt{make} in \texttt{/mpi\_s}: \\
\begin{lstlisting}
$ cd ../mpi_s/
$ make 
\end{lstlisting}
This creates dummy functions for calls to the MPI-library,
which is required for serial compilation.
\item Edit \texttt{Makefile} in \texttt{/src}. To switch off 
the assertions set \texttt{OPTS = -DNP\_D3Q6 -DNDEBUG}. This is
important for production runs (see also comment in \texttt{Makefile}): \\
\begin{lstlisting}
$ cd ../src/
$ vim Makefile
...

@@

OPTS = -DNP_D3Q6 -DNDEBUG
LIBS = -L../targetDP -ltarget -lm
INC = -I. -I ../targetDP
...
\end{lstlisting}
\item To compile the code in serial issue: \\
\begin{lstlisting}
$ make serial
\end{lstlisting} 
To compile the parallel version of the code issue: \\
\begin{lstlisting}
$ make mpi
\end{lstlisting} 
This creates the executable file \texttt{Ludwig.exe} in \texttt{/src}. 
Before going from a serial to a parallel compilation or vice versa
you should clean the directory by issuing
\begin{lstlisting}
$ make clean
\end{lstlisting}

>>>>

The system requirements are an up-to-date version of a C compiler as well as working installations 
of a text editor like \texttt{vi} or \texttt{emacs}, the visualisation tool
ParaView \\
(\hyperref[ParaView]{https://www.paraview.org}), the MPI-library 
for compiling the parallel version of the code and \texttt{gnuplot} for simple visualisations.

@@

The \texttt{Ludwig} project is currently hosted at our central repository at GitHub:\\
(\hyperref[GitHubLudwig]{https://github.com/ludwig-cf/ludwig}).
\smallskip

You can either download or clone the repository. 
The following command clones the repository and requires a working 
git installation on your local system:

@@

\begin{lstlisting}
$ ls ludwig
LICENSE Makefile.mk config docs mpi_s src target tests util version.h
\end{lstlisting}

A full documentation of \texttt{Ludwig} can be obtained by running 
the command \texttt{make pdf} twice in the \texttt{/docs} directory:

\begin{lstlisting}[style=terminalverbatim]
$ cd /ludwig/docs/
$ make pdf
\end{lstlisting}

@@

\begin{enumerate}
\item Go to the top level directory \texttt{/ludwig}: \\
\begin{lstlisting}
$ cd ludwig
\end{lstlisting}
\item Copy the configuration file \texttt{lunix-gcc-default.mk} to top level directory 
and rename to \texttt{config.mk} (both steps exectued below at once): \\
\begin{lstlisting}
$ cp config/lunix-gcc-default.mk ./config.mk
\end{lstlisting}
Note that you may need to specify the path to the MPI wrapper compiler in this file.
\item Issue \texttt{make} in \texttt{/target}: \\
\begin{lstlisting}
$ cd target/
$ make 
\end{lstlisting}
This creates dummy functions for calls to the targetDP library.
\item Issue \texttt{make} in \texttt{/mpi\_s}: \\
\begin{lstlisting}
$ cd ../mpi_s/
$ make 
\end{lstlisting}
This creates dummy functions for calls to the MPI-library,
which is required for serial compilation.
\item Edit \texttt{Makefile} in \texttt{/src}. To switch off 
the assertions set \texttt{OPTS = -DNP\_D3Q6 -DNDEBUG}. This is
important for production runs (see also comment in \texttt{Makefile}): \\
\begin{lstlisting}
$ cd ../src/
$ vim Makefile
...

@@

OPTS = -DNP_D3Q6 -DNDEBUG
LIBS = -L../target -ltarget -lm
INC = -I. -I ../target
...
\end{lstlisting}
\item To compile the code in serial issue: \\
\begin{lstlisting}
$ make serial
\end{lstlisting} 
To compile the parallel version of the code issue: \\
\begin{lstlisting}
$ make mpi
\end{lstlisting} 
This creates the executable file \texttt{Ludwig.exe} in \texttt{/src}. 
Before going from a serial to a parallel compilation or vice versa
you should clean the directory by issuing
\begin{lstlisting}
$ make clean
\end{lstlisting}
